Process Public .Rhistory Files
========================================================

Inspired by [Aldo Cortesi's post](http://corte.si/posts/hacks/github-shhistory/index.html) documenting his exploration of public shell history files posted to GitHub, I was curious if there were any such `.Rhistory` files. For the uninitiated, `.Rhistory` files 


Load Data
---------

```{r, warning=FALSE, cache=TRUE, echo=FALSE}
files <- list()
for (file in list.files("fetched")){
  thisFile <- readLines(paste("fetched/", file, sep=""))
  files[[file]] <- thisFile
}

#Remove empty files
files <- files[sapply(files, length) > 0]
```

This leaves us with a total of `r length(files)` non-empty files, totalling `r sum(sapply(files, length))` commands entered into R.

First, I was curious about the length of these files.

```{r, fig.width=8, fig.height=2, message=FALSE, echo=FALSE}
library(ggplot2)
qplot(sapply(files, length), geom="histogram") + xlab(".Rhistory Length (LOC)")
```

It seems that many of these files represent *very* brief (and likely unpleasant) interaction with R. For instance:

    exit
    exit
    ls
    exit
    
(if you're out there, you were likely looking for the '`q()`' command). Others represent quite extensive projects; the maximum was `r max(sapply(files, length))` lines long.

Package Usage
-------------

More interesting to me was *how* these users were using R -- what the details contained in these history files represent in terms of the user's interaction with R. For starters, which packages were the users using? We can identify packages loaded via the `library()` or `require()` functions.

```{r, cache=TRUE, echo=FALSE}
library(stringr)
flatFile <- as.character(unlist(files))

libraried <- str_match_all(flatFile, "library\\(\"?([^\\,\")]*)\"?[,\\)]")
libraried <- libraried[sapply(libraried, length) > 0]
libraried <- as.character(sapply(libraried, function(x){ return(x[,2]) } ))

required <- str_match_all(flatFile, "require\\(\"?([^\\,\")]*)\"?[,\\)]")
required <- required[sapply(required, length) > 0]
required <- as.character(sapply(required, function(x){ return(x[,2]) } ))

loaded <- c(libraried, required)
```

There were `r length(loaded)` such calls to load packages in the scripts. The top 10 packages loaded in this set were:

Package Name | Count
------------ | ------
```{r, echo=FALSE, results='asis'}
topPkg <- head(sort(table(loaded), decreasing=TRUE),n=10)
cat(paste(names(topPkg), as.integer(topPkg), collapse="\n", sep=" | "))
```

(Of course it's likely worth noting the selection bias from examining only R commands which were included in GitHub projects. I would imagine that the usage for `devtools`, for instance, is certainly higher among GitHub projects.)

Function Use
------------

I'm also curious which functions were most widely executed. We can get a rough identification of most function names by looking for a sequence of valid characters followed by an `(` symbol.

```{r, echo=FALSE, cache=TRUE}
functions <- str_match_all(flatFile, "([[:alnum:]\\._]+)[[:space:]]*\\(")
functions <- functions[sapply(functions, length) > 0]
functions <- as.character(sapply(functions, function(x){ return(x[,2]) } ))
```

This gives us a total of `r length(functions)` function calls. The 10 most popular functions executed in this set were:

Function Name | Count
------------ | ------
```{r, echo=FALSE, results='asis'}
topFun <- head(sort(table(functions), decreasing=TRUE),n=10)
cat(paste(names(topFun), as.integer(topFun), collapse="\n", sep=" | "))
```






