Analysis of Public .Rhistory Files
========================================================

GitHub recently launched a more powerful search feature which has been used on more than one occasion to [identify sensitive files that may be hosted in a public GitHub repository](http://it.slashdot.org/story/13/01/25/132203/github-kills-search-after-hundreds-of-private-keys-exposed). When used innocently, there are all sorts of fun things you can find with this search feature.

Inspired by [Aldo Cortesi's post](http://corte.si/posts/hacks/github-shhistory/index.html) documenting his exploration of public shell history files posted to GitHub, I was curious if there were any such `.Rhistory` files. For the uninitiated, `.Rhistory` files are just logs of commands entered into the interactive console during an R session. Some recent IDEs, such as RStudio, automatically create these files as you work. By default, these files would be excluded from a Git repository, but users could, for whatever reason, choose to include their `.Rhistory` files in the repository.

Using this search function, combined with the [Python script](https://github.com/cortesi/ghrabber) Mr. Cortesi had put together to download the files associated with a GitHub search, I was able to download 638 `.Rhistory` files from public GitHub repositories (excluding forks). What follows is an exploration of those files.

Load Data
---------

```{r, warning=FALSE, cache=TRUE, echo=FALSE}
files <- list()
for (file in list.files("fetched")){
  thisFile <- readLines(paste("fetched/", file, sep=""))
  files[[file]] <- thisFile
}

#Remove empty files
files <- files[sapply(files, length) > 0]
```

Trimming out the 0-line `.Rhistory` files leaves us with a total of `r length(files)` non-empty files, totaling `r sum(sapply(files, length))` commands entered into R.

First, I was curious about the length of these files.

```{r, fig.width=8, fig.height=2, message=FALSE, echo=FALSE}
library(ggplot2)
qplot(sapply(files, length), geom="histogram") + xlab(".Rhistory Length (LOC)")
```

It seems that many of these files represent *very* brief (and likely unpleasant) interaction with R. For instance:

    exit
    exit
    ls
    exit
    
(if you're out there, you were likely looking for the '`q()`' command). Others represent quite extensive projects; the maximum was `r max(sapply(files, length))` lines long.

Package Usage
-------------

More interesting to me was *how* these users were using R -- what the details contained in these history files represent in terms of the user's interaction with R. For starters, which packages were the users using? We can identify packages loaded via the `library()` or `require()` functions.

```{r, cache=TRUE, echo=FALSE}
library(stringr)
flatFile <- as.character(unlist(files))

libraried <- str_match_all(flatFile, "library\\(\"?([^\\,\")]*)\"?[,\\)]")
libraried <- libraried[sapply(libraried, length) > 0]
libraried <- as.character(sapply(libraried, function(x){ return(x[,2]) } ))

required <- str_match_all(flatFile, "require\\(\"?([^\\,\")]*)\"?[,\\)]")
required <- required[sapply(required, length) > 0]
required <- as.character(sapply(required, function(x){ return(x[,2]) } ))

loaded <- c(libraried, required)
```

There were `r length(loaded)` such calls to load packages in the scripts. The top 10 packages loaded in this set were:

Package Name | Count
------------ | ------
```{r, echo=FALSE, results='asis'}
topPkg <- head(sort(table(loaded), decreasing=TRUE),n=10)
cat(paste(paste("`", names(topPkg), "`", sep="") , as.integer(topPkg), collapse="\n", sep=" | "))
```

(Of course it's likely worth noting the selection bias from examining only R commands which were included in GitHub projects. I would imagine that the usage for `devtools`, for instance, is certainly inflated among GitHub projects over the general populace.)

Function Use
------------

I was also curious which functions were most widely executed. We can get a rough identification of most function names by looking for a sequence of valid characters followed by an `(` symbol.

```{r, echo=FALSE, cache=TRUE}
functions <- str_match_all(flatFile, "([[:alnum:]\\._]+)[[:space:]]*\\(")
functions <- functions[sapply(functions, length) > 0]
functions <- as.character(sapply(functions, function(x){ return(x[,2]) } ))
```

This gives us a total of `r length(functions)` function calls of `r length(unique(functions))` unique function names. The 20 most popular functions executed in this set were:

Function Name | Count
------------ | ------
```{r, echo=FALSE, results='asis'}
topFun <- head(sort(table(functions), decreasing=TRUE),n=20)
cat(paste(paste("`", names(topFun), "`", sep="") , as.integer(topFun), collapse="\n", sep=" | "))
```

It should also be possible to identify for which functions the help/manual pages were viewed by identifying lines beginning with a "?" or arguments inside of a call to `help()`.

```{r, echo=FALSE, cache=TRUE}
help <- str_match_all(flatFile, "help\\(\"?([^,^\\^\")]+)\"?,?.*\\)")
help <- help[sapply(help, length) > 0]
help <- as.character(sapply(help, function(x){ return(x[,2]) } ))
#remove calls for package-wide help
help <- help[!grepl("package[[:space:]]?=", help)]

ques <- str_match_all(flatFile, "^\\?[[:space:]]*([[:alnum:]\\._]+)" )
ques <- ques[sapply(ques, length) > 0]
ques <- as.character(sapply(ques, function(x){ return(x[,2]) } ))

qs <- c(help, ques)
```

I can identify `r length(qs)` requests for help on `r length(unique(qs))` different function names. The top 10 most prevalent functions for which users request help follow.

Function Name | Count
------------ | ------
```{r, echo=FALSE, results='asis'}
topHelp <- head(sort(table(qs), decreasing=TRUE),n=10)
cat(paste(paste("`", names(topHelp), "`", sep="") , as.integer(topHelp), collapse="\n", sep=" | "))
```

Conclusion
----------

Of course, there are all sorts of different types of analysis one could perform on this dataset. Post any suggestions you have in the comments; I imagine there's at least one more post of interesting finds in this data.
